Relative Stable Maps in Genus One via Central Alignments

### 2019-01-31 GenusOneRelativeMaps.tex
minor changes
Fixed proof that M_v is always integral (previous proof was wrong)

\begin{lemma} Given a combinatorial type $\Delta$ indexing a stratum of $\Mcal$ and a vertex $v$ of the associated graph, consider the corresponding moduli space of punctured stable maps $\ol\Mcal_v$. As in the above discussion, define for each adjacent flag $E \ni v$ the log stack $\widetilde\Mcal_v^E$ by pulling back the log structure of the punctured curve along the marking section. Consider the fibre product \emph{in the category of coherent log stacks}:
\begin{equation*} \widetilde\Mcal^{\operatorname{not-fs}}_v = \widetilde{\Mcal}_v^{E_1} \times_{\ol\Mcal_v} \ldots \times_{\ol\Mcal_v} \widetilde{\Mcal}_v^{E_k}. \end{equation*}
Then $\widetilde\Mcal^{\operatorname{not-fs}}_v$ is integral (and consequently, the map $\widetilde\Mcal_v \to \widetilde\Mcal^{\operatorname{not-fs}}_v$ is finite).
\end{lemma}
\begin{proof} Locally we have a chart for $\ol\Mcal_v$ given by $Q$ (the minimal monoid as constructed, for instance, in \cite[Construction 1.16]{GrossSiebertLog}) and charts for $\widetilde\Mcal_v^{E_i}$ given by $Q_{E_i}$ where $Q_{E_i} = Q \oplus \N$ if $E_i$ corresponds to an ordinary marked point, and $Q_{E_i}$ is the submonoid of $Q \oplus \Z$ generated by $Q \oplus \N$ and $(\ol\varphi_{\eta_i}(1),u_{E_i})$ if $E_i$ corresponds to a puncturing (so that $u_{E_i} \in \Z_{<0}$). In either case, we have $Q_{E_i} \subseteq Q \oplus \Z \subseteq Q^{\operatorname{gp}} \oplus \Z$ (since $Q$ is integral) and so $Q_{E_i}$ is integral (recall that a monoid is integral if and only if it admits an injective homomorphism into a group). In this notation, $\widetilde\Mcal_v^{\operatorname{not-fs}}$ admits a chart given by the monoid:
\begin{equation*} P = Q_{E_1} \oplus_Q Q_{E_2} \oplus_Q \ldots \oplus_Q Q_{E_k}.\end{equation*}
If we consider the map $D \colon Q^{r-1} \to Q^r$ given by the $(r-1) \times r$ matrix
\begin{equation*} D= \begin{pmatrix} 1 & 1 & 0 & 0 & \ldots & 0 \\
0 & 1 & 1 & 0 & \ldots & 0 \\
\, & \, & \, &\vdots & \, & \, \\
0 & 0 & \ldots & 0 & 1 & 1 \end{pmatrix}\end{equation*}
then the monoid $P$ can be expressed as a cokernel
\begin{equation*} Q^{r-1} \to \oplus_{i=1}^r Q_{E_i} \to P \end{equation*}
where the first map is obtained by composing $D$ with the natural inclusion $Q^r \to \oplus_{i=1}^r Q_{E_i}$. We may now form a diagram
\bcd
Q^{r-1} \ar[d,"\Id"] \ar[r] & \oplus_{i=1}^r Q_{E_i} \ar[r] \ar[d,hook] & P \ar[d] \\
Q^{r-1} \ar[r] & Q^r \oplus \Z^r \ar[r] & S \oplus \Z^r
\ecd
where both rows are cokernel diagrams and the map $P \to S\oplus\Z^r$ is induced by the universal property. Here $S$ is the cokernel of $D$ and is given by:
\begin{equation*} S = \begin{cases} Q \qquad \text{if } r=1 \\ Q^{\gp} \qquad \text{if } r \geq 2 \end{cases}.\end{equation*}
In any case, $S$ is integral and the map $P \to S \oplus \Z^r$ is injective because the middle vertical maps in injective. Hence we conclude that $P$ is integral, as required. \end{proof}

The previous lemma holds for arbitrary moduli spaces of punctured maps. The next lemma, on the other hand, is special to our setting.

\begin{lemma} Consider the fibre product in the category of coherent log stacks:
\bcd
\Scal^{\operatorname{not-fs}} \ar[r] \ar[d] \ar[rd,phantom,"\square"] & \prod_v \widetilde\Mcal_v \ar[d] \\
\prod_{E} Z_E \ar[r] & \prod_{E} Z_E \times Z_E.
\ecd
Then $\Scal^{\operatorname{not-fs}}$ is integral.\end{lemma}

>>>>

\begin{lemma} Given a combinatorial type $\Delta$ indexing a stratum of $\Mcal$ and a vertex $v$ of the associated graph, consider the corresponding moduli space of punctured stable maps $\ol\Mcal_v$. As in the above discussion, define for each adjacent flag $E \ni v$ the log stack $\widetilde\Mcal_v^E$ by pulling back the log structure of the punctured curve along the marking section. Consider the fibre product \emph{in the category of coherent log stacks}:
\begin{equation*} \widetilde\Mcal^{\operatorname{not-fs}}_v = \widetilde{\Mcal}_v^{E_1} \times_{\ol\Mcal_v} \ldots \times_{\ol\Mcal_v} \widetilde{\Mcal}_v^{E_k}. \end{equation*}
Then $\widetilde\Mcal^{\operatorname{not-fs}}_v$ is integral (and consequently, the map $\widetilde\Mcal_v \to \widetilde\Mcal^{\operatorname{not-fs}}_v$ is finite).
\end{lemma}
\begin{proof} Locally we have a chart for $\ol\Mcal_v$ given by $Q$ (the minimal monoid as constructed, for instance, in \cite[Construction 1.16]{GrossSiebertLog}) and charts for $\widetilde\Mcal_v^{E_i}$ given by $Q_{E_i}$, where $Q_{E_i} = Q \oplus \N$ if $E_i$ corresponds to an ordinary marked point, and $Q_{E_i}$ is the submonoid of $Q \oplus \Z$ generated by $Q \oplus \N$ and $(\ol\varphi_{\eta_i}(1),u_{E_i})$ if $E_i$ corresponds to a puncturing (so that $u_{E_i} \in \Z_{<0}$). In either case, we have $Q_{E_i} \subseteq Q \oplus \Z \subseteq Q^{\operatorname{gp}} \oplus \Z$ (since $Q$ is integral) and so $Q_{E_i}$ is integral (recall that a monoid is integral if and only if it admits an injective homomorphism into a group). In this notation, $\widetilde\Mcal_v^{\operatorname{not-fs}}$ admits a chart given by the monoid:
\begin{equation*} P = Q_{E_1} \oplus_Q Q_{E_2} \oplus_Q \ldots \oplus_Q Q_{E_k}.\end{equation*}
Suppose without loss of generality that $E_1,\ldots,E_r$ are punctures and $E_{r+1},\ldots,E_k$ are ordinary marked points; then $P$ is equal to the submonoid of
\begin{equation*} Q \times \Z^r \times \N^{k-r} \end{equation*}
generated by $Q \times \N^k$ and $(\ol\varphi_{\eta_i}(1),u_{E_i}\cdot e_i,0)$ where $e_i \in \Z^r$ is the $i$th basis element and $0 \in \N^r$. In particular, $P$ admits an injective homomorphism into the group
\begin{equation*} Q^{\gp} \times \Z^r \times \Z^{k-r} \end{equation*}
and hence is integral.\end{proof}

\footnote{(Navid) Is this a good point to calculate the degree of $\widetilde\Mcal_v \to \widetilde\Mcal_v^{\operatorname{not-fs}}$?}The previous lemma holds for arbitrary moduli spaces of punctured maps. The next lemma, on the other hand, is special to our setting.

\begin{lemma} Consider the fibre product in the category of coherent log stacks:
\bcd
\Scal^{\operatorname{not-fs}} \ar[r] \ar[d] \ar[rd,phantom,"\square"] & \prod_v \widetilde\Mcal_v \ar[d] \\
\prod_{E} Z_E \ar[r] & \prod_{E} Z_E \times Z_E.
\ecd
Then $\Scal^{\operatorname{not-fs}}$ is integral.\end{lemma}
\begin{proof}\end{proof}
